Proofs for ``Geometry of Local Spectral Clustering''

### 2020-12-30 jmlr/revision/proofs.tex
Completed fixed graph proofs.

%%% Theorem environments
\newtheorem{theorem}{Theorem}
\newtheorem{conjecture}{Conjecture}
\newtheorem{lemma}{Lemma}
\newtheorem{example}{Example}
\newtheorem{corollary}{Corollary}
\newtheorem{proposition}{Proposition}
\newtheorem{assumption}{Assumption}
\newtheorem{remark}{Remark}

@@

The proofs of our major theorems all consist of at most three modular parts.
\begin{enumerate}
	\item~\textbf{Fixed graph results.} Analysis which holds with respect to an arbitrary graph $G$, and are stated with respect to functionals (i.e. normalized cut, conductance, and local spread) of $G$;
	\item~\textbf{Finite-sample bounds.} For the specific choice of $G = G_{n,r}$, we relate the aforementioned functionals to their population analogues. This proves~\textcolor{red}{(?)} 
	\item~\textbf{Population functionals.} (In the case of density clustering only.) When the candidate cluster is a $\lambda$-density cluster, we bound the population functionals by a function of $\lambda$, as well as the other relevant parameters introduced in Section~\textcolor{red}{(?)}. This proves~\textcolor{red}{(?)}
\end{enumerate}
The first three sections of this appendix will correspond to each of these three parts. In Section~\ref{sec:pf_major_theorems}, we will combine these parts to prove the major theorems of our main text, \textcolor{red}{(...)}. Finally, in Section~\ref{sec:appr_misclassification_error} we derive upper bounds for the aPPR vector, and in Section~\ref{sec:experimental_details} we give relevant details regarding our experiments.

@@

\paragraph{Notation.}
The graph $G = (V,E)$ is an undirected and connected but otherwise arbitrary graph, defined over vertices $V = \{1,\ldots,n\}$ with $m = |E|$ total edges. The adjacency matrix of $G$ is $A$, the degree matrix is $D$, and the lazy random walk matrix over $G$ is $W = (I + D^{-1}A)/2$. If the lazy random walk originates at a node $v$, the distribution of the lazy random walk $q_v^{(t)} := q(v,t;G)$ after $t$ steps is $q_v^{(t)} := e_v W^t$, with stationary distribution $\pi := \lim_{t \to \infty} q_v^{(t)}$ with entries $\pi(u) = \deg(u;G)/\vol(u;G)$.

The PPR vector $p_v = p(v,\alpha;G)$ is defined as the solution to~\textcolor{red}{(?)}

Given a distribution $q$ (by distribution we mean a vector with non-negative entries; for instance, $q = q_v^{(t)}$ for $t \in \mathbb{N}$, $q = p_v$, or $q = \pi$) and $\beta \in (0,1)$, the $\beta$-sweep cut of $q$ is
\begin{equation*}
S_{\beta}(q) = \set{u: \frac{q(u)}{\deg(u;G)} > \beta};
\end{equation*} 
in the special case where $q = p_v$ we write $S_{\beta,v}$ for $S_{\beta}(p_v)$. The argument of $S_{\beta}(\cdot)$ will usually be clear from context, in which case we will drop it and simply write $S_{\beta}$. For $j = 1,\ldots,n$, let $\beta_j$ be the smallest value of $\beta \in (0,1)$ such that the sweep cut $S_{\beta_j}$ contains at least $j$ vertices. For notational ease, we will write $S_j := S_{\beta_j}$, and $S_0 = \emptyset$. We adopt the

We now introduce a function $h_q(\cdot): [0,2m] \to [0,1]$ to measure the extent to which a distribution $q$ is mixed. To do so, we first define a piecewise linear function $q[\cdot]: [0,2m] \to [0,1]$. Letting $q(S) := \sum_{u \in S} q(u)$, we take $q[\vol(S_j)] = q(S_j)$ for each sweep cut $S_j$, and then extend $q[\cdot]$ by piecewise linear interpolation to be defined everywhere on its domain. Then the mixedness of $q$ is measured by
\begin{equation*}
h_q(k) := q[k] - \frac{k}{2m}.
\end{equation*}
Note that $h_{\pi} = 0$. By construction both $q[\cdot]$ and $h_q(\cdot)$ are concave functions, which will be an important fact later on.

% For a connected subset of the vertices $C \subseteq V$, the subgraph induced by $C$ is $G[C] = (C, E \cap (C \times C))$. The lazy random walk matrix over $C$ is $W_C = (I_C + D^{-1}_C A_C)/2$ (where for a matrix $M \in \Reals^{n \times n}$ we write $M_C$ for the submatrix of $M$ induced by $C$), the vector  
%\begin{equation}
%\label{eqn:lazy_random_walk}
%q_{v}^{(t)}(u) = e_vW_C^t e_u
%\end{equation}
%denotes the probability that after $t$-steps the lazy random walk originating at $v$ arrives at $u$. The stationary distribution of this random walk is $\pi_C$.

The conductance of the vertex set $V$ is abbreviated as $\Psi(G) := \Psi(V;G)$, and likewise for the local spread $s(G) := s(V;G)$. Finally, for convenience we introduce the following functionals:
\begin{equation*}
\begin{aligned}
& d_{\max}(C; G) := \max_{u \in C} \deg(u; G), && d_{\min}(C; G) := \min_{u \in C} \deg(u;G) \\
& d_{\max}(G) := d_{\max}(V;G),~~ && d_{\min}(G) := d_{\min}(V; G)
\end{aligned}
\end{equation*}

In the following subsections we establish: (1) an upper bound on the misclassification error of PPR in terms of $\alpha$ and $\Phi(C;G)$ (Lemma~\ref{lem:zhu}); (2) upper bounds on the mixedness of $q_v^{(t)}$ (as a function of $t$) and $p_v$ (as a function of $\alpha$), which will be helpful in the proofs of Proposition~\ref{prop:pointwise_mixing_time} and Theorem~\ref{thm:ppr_lb}; (3) an upper bound on $\tau_{\infty}(G)$ in terms of $\Psi(G)$ and $s(G)$ (Proposition~\ref{prop:pointwise_mixing_time}); (4) an upper bound on the normalized cut $\Phi(\wh{C};G)$ in terms of $\Phi(C;G)$, to be used later in the proof of Theorem~\ref{thm:ppr_lb} (negative example); and (5) a uniform bound on the perturbations of the PPR vector, to be used later in the proof of Theorem~\ref{thm:density_cluster_consistent_recovery} (consistency of PPR). 
\subsection{Proof of Lemma~\ref{lem:zhu}}
\label{subsec:pf_lem_zhu}
\textcolor{red}{(TODO)}

\subsection{Mixedness of lazy random walk and PPR vectors}
In this subsection, we give upper bounds on $h^{(t)} := h_{q_v^{(t)}}$ and $h^{(\alpha)} := h_{p_v}$. Although similar bounds exist in the literature (see in particular Theorem~1.1 of \citep{lovasz1990} and Theorem~3 of~\citep{andersen2006}), we could not find precisely the results we needed, and so for completeness we state and prove these results ourselves.

\begin{theorem}
	\label{thm:mixing_time_rw}
	For any $k \in [d_{\max}(G), 2m - d_{\max}(G)]$ and $t \geq 3$,
	\begin{equation}
	\label{eqn:mixing_time_rw_1}
	h^{(t)}(k) \leq \frac{1}{8} + \frac{d_{\max}(G)}{2d_{\min}(G)^2} + \frac{\sqrt{m}}{d_{\min}(G)^2} \sqrt{\wb{k}} \Bigl(1 - \frac{\Psi(G)^2}{8}\Bigr)^{t - 3}.
	\end{equation}
	For any $k \in [0, d_{\max}(G)) \cup (2m - d_{\max}(G),2m]$,
	\begin{equation}
	\label{eqn:mixing_time_rw_2}
	h^{(t)}(k) \leq \frac{1}{8} + \frac{d_{\max}(G)}{2d_{\min}(G)^2}.
	\end{equation}
\end{theorem}

\begin{theorem}
	\label{thm:mixing_time_PPR}
	Let $\phi$ be any constant in $[0,1]$. Either the following bound holds for any $t \in \mathbb{N}$ and any $k \in [d_{\max}(G),2m - d_{\max}(G)]$:
	\begin{equation*}
	h^{(\alpha)}(k) \leq \alpha t + \frac{2\alpha}{1 + \alpha} + \frac{d(v)}{2d_{\min}^2} + \frac{\sqrt{m}}{d_{\min}^2} \cdot \sqrt{\overline{k}} \left(1 - \frac{\phi^2}{8}\right)^{t},
	\end{equation*}
	or there exists some sweep cut $S_j$ of $p_v$ such that $\Phi(S_j;G) \leq \phi$.
\end{theorem}

The proofs of these upper bounds will be similar to each other (in places word-for-word alike), and will follow a similar approach and use similar notation to that of \citep{lovasz1990,andersen2006}. For a given $0 \leq K_0 \leq m$ and a function $h: [0,2m] \to [0,1]$, let 
\begin{equation*}
L_{K_0}(k;h) = \frac{2m - K_0 - k}{2m - 2K_0}h(K_0) + \frac{k - K_0}{2m - 2K_0}h(2m - K_0)
\end{equation*}
be the linear interpolant of $h(K_0)$ and $h(2m - K_0)$, and additionally let
\begin{equation*}
C(K_0;h) := \max\set{\frac{h(k) - L_{K_0}(k;h)}{\sqrt{\overline{k}}}: K_0 \leq k \leq  2m - K_0}.
\end{equation*}
where we use the notation $\overline{k} := \min\{k, 2m - k\}$, and set $0/0$ equal to $0$. Our first pair of Lemmas upper bound $h^{(t)}$ and $h^{(\alpha)}$ as a function of $L_{K_0}$ and $C(K_0)$. Lemma~\ref{lem:mixing_random_walk} implies that if $t$ is large relative to $\Psi(G)$, then $h^{(t)}(\cdot)$ must be small.

@@

\begin{lemma}[\textbf{c.f Theorem~3 of \citep{andersen2006}}]
	\label{lem:mixing_time_PPR}
	Let $\phi \in [0,1]$. Either the following bound holds for any $t \in \mathbb{N}$, any $K_0 \in [0,m]$, and any $k \in [K_0,2m - K_0]$:
	\begin{equation}
	\label{eqn:mixing_time_PPR}
	h^{(\alpha)}(k) \leq \alpha t + L_{K_0}(k; h^{(\alpha)}) + C(K_0;h^{(\alpha)})\sqrt{\overline{k}}\left(1 - \frac{\phi^2}{8}\right)^t
	\end{equation}
	or else there exists some sweep cut $S_{j}$ of $p_v$ such that $\Phi(S_j;G) \leq \phi$.
\end{lemma}

>>>>

%%% Theorem environments
\newtheorem{theorem}{Theorem}[section]
\newtheorem{conjecture}{Conjecture}[section]
\newtheorem{lemma}{Lemma}[section]
\newtheorem{example}{Example}[section]
\newtheorem{corollary}{Corollary}[section]
\newtheorem{proposition}{Proposition}[section]
\newtheorem{assumption}{Assumption}[section]
\newtheorem{remark}{Remark}[section]

@@

The proofs of our major theorems all consist of at most three modular parts.
\begin{enumerate}
	\item~\textbf{Fixed graph results.} Results which hold with respect to an arbitrary graph $G$, and are stated with respect to functionals (i.e. normalized cut, conductance, and local spread) of $G$;
	\item~\textbf{Finite-sample bounds.} For the specific choice of $G = G_{n,r}$, we relate the aforementioned functionals to their population analogues. This proves~\textcolor{red}{(?)} 
	\item~\textbf{Population functionals.} (In the case of density clustering only.) When the candidate cluster is a $\lambda$-density cluster, we bound the population functionals by a function of $\lambda$, as well as the other relevant parameters introduced in Section~\textcolor{red}{(?)}. This proves~\textcolor{red}{(?)}
\end{enumerate}
The first three sections of this appendix will correspond to each of these three parts. In Section~\ref{sec:pf_major_theorems}, we will combine these parts to prove the major theorems of our main text, \textcolor{red}{(...)}. Finally, in Section~\ref{sec:appr_misclassification_error} we derive upper bounds for the aPPR vector, and in Section~\ref{sec:experimental_details} we give relevant details regarding our experiments.

@@

\paragraph{Notation.}
The graph $G = (V,E)$ is an undirected and connected but otherwise arbitrary graph, defined over vertices $V = \{1,\ldots,n\}$ with $m = |E|$ total edges. The adjacency matrix of $G$ is $A$, the degree matrix is $D$, and the lazy random walk matrix over $G$ is $W = (I + D^{-1}A)/2$. If the lazy random walk originates at a node $v$, the distribution of the lazy random walk $q_v^{(t)} := q(v,t;G)$ after $t$ steps is $q_v^{(t)} := e_v W^t$, with stationary distribution $\pi := \pi(G) := \lim_{t \to \infty} q_v^{(t)}$ with entries $\pi(u) = \deg(u;G)/\vol(u;G)$.

For a starting distribution $s$ (by distribution we mean a vector with non-negative entries), the PPR vector $p_s = p(s,\alpha;G)$ is the solution of
\begin{equation}
\label{eqn:ppr}
p_s = \alpha s + (1 - \alpha) p_s W.
\end{equation}
When $s = e_v$, we write $p_v := p_{e_v}$. It is easy to check that $p_s = \alpha \sum_{t = 0}^{\infty} (1 - \alpha)^t q_s^{(t)}$.  Note that $s$ need not be a probability distribution (i.e. its entries need not sum to $1$) to make sense of~\eqref{eqn:ppr}.

Given a distribution $q$ (for instance, $q = q_v^{(t)}$ for $t \in \mathbb{N}$, $q = p_v$, or $q = \pi$) and $\beta \in (0,1)$, the $\beta$-sweep cut of $q$ is
\begin{equation*}
S_{\beta}(q) = \set{u: \frac{q(u)}{\deg(u;G)} > \beta};
\end{equation*} 
in the special case where $q = p_v$ we write $S_{\beta,v}$ for $S_{\beta}(p_v)$. The argument of $S_{\beta}(\cdot)$ will usually be clear from context, in which case we will drop it and simply write $S_{\beta}$. For $j = 1,\ldots,n$, let $\beta_j$ be the smallest value of $\beta \in (0,1)$ such that the sweep cut $S_{\beta_j}$ contains at least $j$ vertices. For notational ease, we will write $S_j := S_{\beta_j}$, and $S_0 = \emptyset$.

We now introduce the~\emph{Lovasz-Simonovits curve} $h_q(\cdot): [0,2m] \to [0,1]$ to measure the extent to which a distribution $q$ is mixed. To do so, we first define a piecewise linear function $q[\cdot]: [0,2m] \to [0,1]$. Letting $q(S) := \sum_{u \in S} q(u)$, we take $q[\vol(S_j)] = q(S_j)$ for each sweep cut $S_j$, and then extend $q[\cdot]$ by piecewise linear interpolation to be defined everywhere on its domain. Then the mixedness of $q$ is measured by
\begin{equation*}
h_q(k) := q[k] - \frac{k}{2m}.
\end{equation*}
The Lovasz-Simonovits curve is a non-negative function, with $h_q(0) = h_q(2m) = 0$. The stationary distribution $\pi$ is mixed, i.e. $h_{\pi}(k) = 0$ for all $k \in [0,2m]$. Finally, both $q[\cdot]$ and $h_q(\cdot)$ are concave functions, which will be an important fact later on.

The conductance of $V$ is abbreviated as $\Psi(G) := \Psi(V;G)$, and likewise for the local spread $s(G) := s(V;G)$. Finally, for convenience we introduce the following functionals:
\begin{equation*}
\begin{aligned}
& d_{\max}(C; G) := \max_{u \in C} \deg(u; G), && d_{\min}(C; G) := \min_{u \in C} \deg(u;G) \\
& d_{\max}(G) := d_{\max}(V;G),~~ && d_{\min}(G) := d_{\min}(V; G)
\end{aligned}
\end{equation*}
We note that $d_{\min}(G)^2 \leq d_{\min}(G) \cdot n \leq \vol(G) \leq n \cdot d_{\max}(G)$, and that for any $S \subseteq V$, $|S| \cdot d_{\min}(G) \leq \vol(S;G)$ (where $|S|$ is the cardinality of $S$.)

\paragraph{Organization.} In the following subsections we establish: (\ref{subsec:pf_lem_zhu}) an upper bound on the misclassification error of PPR in terms of $\alpha$ and $\Phi(C;G)$ (Lemma~\ref{lem:zhu}); (\ref{subsec:ppr_uniform_bounds}) a uniform bound on the perturbations of the PPR vector, to be used later in the proof of Theorem~\ref{thm:density_cluster_consistent_recovery} (consistency of PPR);  (\ref{subsec:lovasz_simonovits_bounds}) upper bounds on the mixedness of $q_v^{(t)}$ (as a function of $t$) and $p_v$ (as a function of $\alpha$), which will be helpful in the proofs of Proposition~\ref{prop:pointwise_mixing_time} and Theorem~\ref{thm:ppr_lb}; (\ref{subsec:pf_prop_pointwise_mixing_time}) an upper bound on $\tau_{\infty}(G)$ in terms of $\Psi(G)$ and $s(G)$ (Proposition~\ref{prop:pointwise_mixing_time}); and (\ref{subsec:ppr_spectral_partitioning}) an upper bound on the normalized cut $\Phi(\wh{C};G)$ in terms of $\Phi(C;G)$, to be used later in the proof of Theorem~\ref{thm:ppr_lb} (negative example).

\subsection{Proof of Lemma~\ref{lem:zhu}}
\label{subsec:pf_lem_zhu}
For a candidate cluster $C \subseteq V$, we use the tilde-notation $\wt{G} = G[C]$ to refer to the subgraph of $G$ induced by $C$. Similarly we write $\wt{q}_v^{(t)} := q(v,t;\wt{G})$ for the $t$-step distribution of the lazy random walk over $\wt{G}$, $\wt{\pi} = \pi(G[C])$ for the stationary distribution of $\wt{q}_v^{(t)}$ (we will always assume $G[C]$ is connected), and $\wt{p}_v := p(v,\alpha;\wt{G})$ for the PPR vector over $\wt{G}$.

As mentioned in the main text, Lemma~\ref{lem:zhu} is equivalent, up to constants, to Lemma~3.4 in \cite{zhu2013}, and the proof of Lemma~\ref{lem:zhu} proceeds along very similar lines to the proof of that lemma. In fact, we directly use the following three inequalities, derived in that work:
\begin{itemize}
	\item \textbf{(c.f. Lemma 3.2 of \cite{zhu2013})} For any seed node $v \in C$, the PPR vector is lower bounded,
	\begin{equation}
	\label{pf:zhu1}
	\wt{p}_v(u) \geq \frac{3}{4}\bigl(1 - \alpha \cdot \tau_{\infty}(\wt{G})\bigr) \cdot \wt{\pi}(u),~~\textrm{for every $u \in C$.}
	\end{equation}
	\item \textbf{(c.f. Corollary 3.3 of \cite{zhu2013})} For any seed node $v \in C$, there exists a so-called leakage distribution $\ell = \ell(v)$ such that $\mathrm{supp}(\ell) \subseteq C$, $\|\ell\|_1 \leq 2\Phi(C;G)/\alpha$, and 
	\begin{equation}
	\label{pf:zhu2}
	p_v(u) \geq \wt{p}_v(u) - \wt{p}_{\ell}(u),~~\textrm{for every $u \in C$.}
	\end{equation}
	\item \textbf{(c.f. Lemma 3.1 of \cite{zhu2013})} There exists a set $C^g \subset C$ with $\vol(C^g;G) \geq \frac{1}{2}\vol(C;G)$ such that for any seed node $v \in C^g$, the following inequality holds
	\begin{equation}
	\label{pf:zhu3}
	p_v(C^c) \leq 2\frac{\Phi(C;G)}{\alpha}.
	\end{equation}
\end{itemize}
We use~\eqref{pf:zhu1}-\eqref{pf:zhu3} to separately upper bound $\vol(S_{\beta,v} \setminus C;G)$, $\vol(C^{\mathrm{int}} \setminus S_{\beta,v};G)$ and $\vol(C^{\mathrm{bdry}} \setminus S_{\beta,v};G)$; here $C^{\mathrm{int}} \cup C^{\mathrm{bdry}} = C$ is a partition of $C$, with
\begin{equation*}
C^{\mathrm{int}} := \Bigl\{u \in C: \deg(u;\wt{G}) > \bigl(1 - \alpha \cdot \beta \cdot \vol(C;G)\bigr) \deg(u;G) \Bigr\},
\end{equation*}
consisting of those vertices $u \in C$ with sufficient large degree in $\wt{G}$.

First we upper bound $\vol(S_{\beta,v} \setminus C;G)$. Observe that for any $u \in S_{\beta,v} \setminus C$, $p_v(u) > \beta \cdot \deg(u;G)$. Summing up over all such vertices, from~\eqref{pf:zhu3} we conclude that
\begin{equation}
\label{pf:zhu3.5}
\vol(S_{\beta,v} \setminus C; G) \leq \frac{p_v(C^c)}{\beta} \leq 2\frac{\Phi(C;G)}{\beta \cdot \alpha}.
\end{equation}

Next we upper bound $\vol(C^{\mathrm{int}} \setminus S_{\beta,v};G)$. From~\eqref{pf:zhu1} and~\eqref{pf:zhu2} we see that 
\begin{equation*}
p_v(u) \geq \frac{3}{4}\bigl(1 - \alpha \cdot \tau_{\infty}(\wt{G})\bigr) \cdot \wt{\pi}(u) - \wt{p}_{\ell}(u)~~\textrm{for all $u \in C$.}
\end{equation*}
If additionally $u \not\in S_{\beta,v}$ then $p_v(u) \leq \beta \deg(u;G)$, and for all such $u \in C \setminus S_{\beta,v}$,
\begin{equation}
\label{pf:zhu4}
\frac{3}{4}\bigl(1 - \alpha \cdot \tau_{\infty}(\wt{G})\bigr) \cdot \wt{\pi}(u) -  \beta\deg(u;G) \leq \wt{p}_{\ell}(u).
\end{equation}
On the other hand, for any $u \in C^{\mathrm{int}}$ it holds that
\begin{equation*}
\wt{\pi}(u) = \frac{\deg(u;\wt{G})}{\vol(\wt{G})} \geq \frac{\deg(u;\wt{G})}{\vol(G)} \geq \frac{(1 - \alpha \beta \vol(C;G))\deg(u;G)}{\vol(C;G)};
\end{equation*}
by plugging this in to~\eqref{pf:zhu4} we obtain
\begin{equation*}
\biggl(\frac{3(1 - \alpha \beta \vol(C;G))\cdot\bigl(1 - \alpha \tau_{\infty}(\wt{G})\bigr)}{4\vol(C;G)} - \beta\biggr) \cdot \deg(u;G) \leq \wt{p}_{\ell}(u),~~\textrm{for all $u \in C^{\mathrm{int}} \setminus S_{\beta,v}$},
\end{equation*}
and summing over all such $u$ gives
\begin{equation*}
\biggl(\frac{3(1 - \alpha \beta \vol(C;G))\cdot\bigl(1 - \alpha \tau_{\infty}(\wt{G})\bigr)}{4\vol(C;G)} - \beta\biggr) \cdot \vol\bigl(C^{\mathrm{int}} \setminus S_{\beta,v}; G\bigr) \leq \wt{p}_{\ell}\bigl(C^{\mathrm{int}} \setminus S_{\beta,v}\bigr) \leq 2\frac{\Phi(C;G)}{\alpha}.
\end{equation*}
The upper bounds in~\eqref{eqn:zhu_condition} imply
\begin{equation*}
\biggl(\frac{3(1 - \alpha \beta \vol(C;G))\cdot\bigl(1 - \alpha \tau_{\infty}(\wt{G})\bigr)}{4\vol(C;G)} - \beta\biggr) \geq \frac{2}{3}\beta,
\end{equation*}
and we conclude that
\begin{equation}
\label{pf:zhu5}
\vol(C^{\mathrm{int}} \setminus S_{\beta,v}; G) \leq \frac{3\Phi(C;G)}{\alpha\beta}.
\end{equation}

Finally, we upper bound $\vol(C^{\mathrm{bdry}} \setminus S_{\beta,v};G)$. Indeed, for any $u \in C^{\mathrm{bdry}}$,
\begin{equation*}
\frac{1}{\vol(C;G)}\sum_{w \not\in C} \1((u,w) \in E) \geq \alpha \cdot \beta \cdot \deg(u;G)
\end{equation*}
and summing over all such vertices yields
\begin{equation}
\label{pf:zhu6}
\vol(C^{\mathrm{bdry}};G) \leq \frac{1}{\alpha \beta \vol(C;G)}\sum_{\substack{u \in C^{\mathrm{bdry}} \\ w \not\in C}} \1((u,w) \in E) \leq \frac{\Phi(C;G)}{\alpha \cdot \beta}.
\end{equation} 
The claim follows upon summing the upper bounds in~\eqref{pf:zhu3.5}, \eqref{pf:zhu5} and~\eqref{pf:zhu6}.

\subsection{Uniform bounds on PPR}
\label{subsec:ppr_uniform_bounds}
As mentioned in our main text, in order to prove Theorem~\ref{thm:density_cluster_consistent_recovery}, we require a uniform bound on the PPR vector. Actually, we require two such bounds: for a candidate cluster $C \subseteq V$ and an alternative cluster $C' \subseteq V$, we require a lower bound on $p_v(u)$ for all $u \in C$, and an upper bound on $p_v(u')$ for all $u' \in C'$. In Lemma~\ref{lem:ppr_uniform_bound} we establish an upper bound that holds for all vertices $u$ in the interior $C_{o}$ of $C$, and a lower bound holds for all vertices $u'$ in the interior of $C_{o}'$ of $C'$; here
\begin{equation*}
C_{o} = \Bigl\{u \in C: \deg(u,\wt{G}) =  \deg(u;G)\Bigr\},~~\textrm{and}~~C_{o}'= \Bigl\{u \in C': \deg(u,G[C']) =  \deg(u;G)\Bigr\},
\end{equation*}
and we remind the reader that $\wt{G} = G[C]$. 
\begin{lemma}
	\label{lem:ppr_uniform_bound}
	Let $C$ and $C'$ be disjoint subsets of $V$, and suppose that
	\begin{equation*}
	\alpha \leq \frac{1}{2\tau_{\infty}(\wt{G})}.
	\end{equation*}
	Then there exists a set $C^g \subseteq C$ with $\vol(C^g;G) \geq \vol(C;G)/2$ such that for any $v \in C^g$,
	\begin{equation}
	\label{eqn:ppr_uniform_bound_C}
	p_v(u) \geq \frac{3}{8}\wt{\pi}(u) - \frac{2 \Phi(C;G)}{d_{\min}(\wt{G})\cdot \alpha}~~\textrm{for all $u \in C_{o}$}
	\end{equation}
	and
	\begin{equation}
	\label{eqn:ppr_uniform_bound_Cprime}
	p_v(u') \leq \frac{2\Phi(C;G)}{d_{\min}(C';G) \cdot \alpha}~~\textrm{for all $u \in C_{o}'$.}
	\end{equation}
\end{lemma}

\paragraph{``Leakage'' and ``soakage'' vectors.} To prove Lemma~\ref{lem:ppr_uniform_bound}, we will make use of the following explicit representation of the \emph{leakage} distribution $\ell$ from~\eqref{pf:zhu3}, as well as an analogously defined \emph{soakage} distribution $s$:
\begin{equation}
\label{eqn:leakage_soakage}
\begin{aligned}
\ell^{(t)} & := e_v(W \wt{I})^t(I - D^{-1}\wt{D}),~~&& \ell = \sum_{t = 0}^{\infty} (1 - \alpha)^t \ell^{(t)} \\
s^{(t)} & := e_v(W \wt{I})^t W (I - \wt{I}),~~&& s = \sum_{t = 0}^{\infty} (1 - \alpha)^t s^{(t)}.
\end{aligned}
\end{equation}
In the above, $\wt{I} \in \Reals^{n \times n}$ is a diagonal matrix with $I_{uu} = 1$ if $u \in C$ and $0$ otherwise, and $\wt{D}$ is the diagonal matrix with $\wt{D}_{uu} = \deg(u;\wt{G})$ if $u \in C$, and $0$ otherwise.

These quantities admit a natural interpretation in terms of random walks. For $u \in C$, $\ell^{(t)}(u)$ is the probability that a lazy random walk over $G$ originating at $v$ stays within $\wt{G}$ for $t$ steps, arriving at $u$ on the $t$th step, and then ``leaks out'' of $C$ on the $(t + 1)$st step. On the other hand, for $u \not\in C$, $s^{(t)}(u)$ is the probability that a lazy random walk over $G$ originating at $v$ stays within $\wt{G}$ for $t$ steps and is then ``soaked up'' into $u$ on the $(t + 1)$st step. The vectors $\ell$ and $s$ then give the total mass leaked and soaked, respectively, by the PPR vector.

Three properties of $\ell$ and $s$ are worth pointing out. First, $\mathrm{supp}(\ell) \subseteq C \setminus C_o$ and $\mathrm{supp}(s) \subseteq V \setminus C$. Second, $\|\ell^{(t)}\|_1 = \|s^{(t)}\|_1$ for all $t \in \mathbb{N}$, and so $\|\ell\|_1 = \|s\|_1$. Third, for any $u \in V \setminus C$, $p_v(u) = p_s(u)$. The first two properties are immediate. The third property follows by the law of total probability, which implies that
\begin{equation*}
q_v^{(\tau)}(u) = \sum_{t = 0}^{\tau} q_{s^{(t)}}^{(\tau - t)}(u),~~\textrm{for all $u \in V \setminus C$.}
\end{equation*}
or in terms of the PPR vector,
\begin{equation*}
p_v(u) = \alpha \sum_{\tau = 0}^{\infty} (1 - \alpha)^{\tau} q_v^{(\tau)}(u) = \alpha \sum_{\tau = 0}^{\infty} \sum_{t = 0}^{\tau} (1 - \alpha)^{\tau} q_{s^{(t)}}^{(\tau - t)}(u).
\end{equation*}
Substituting $\Delta = \tau + t$ and rearranging gives the claimed property, as
\begin{equation*}
p_v(u) = \alpha \sum_{\tau = 0}^{\infty} \sum_{t = 0}^{\tau} (1 - \alpha)^{\tau} q_{s^{(t)}}^{(\tau - t)}(u) = \sum_{\Delta = 0}^{\infty} \sum_{t = 0}^{\infty} (1 - \alpha)^{\Delta + t} q_{s^{(t)}}^{(\Delta)}(u) = \alpha \sum_{\Delta = 0}^{\infty} (1 - \alpha)^{\Delta} q_s^{(\Delta)}(u) = p_s(u).
\end{equation*}

\begin{proof}[Proof (of Lemma~\ref{lem:ppr_uniform_bound})]
	We first show~\eqref{eqn:ppr_uniform_bound_C}. From~\eqref{pf:zhu3} and~\eqref{pf:zhu2}, we have that
	\begin{equation*}
	p_v(u) \geq \frac{3}{4}\bigl(1 - \alpha \cdot \tau_{\infty}(\wt{G})\bigr) \cdot \wt{\pi}(u) - \wt{p}_{\ell}(u)~~\textrm{for all $u \in C$,}
	\end{equation*}
	where $\ell$ has support $\mathrm{supp}(\ell) \subseteq C$ with $\|\ell\|_1 \leq 2\Phi(C;G)/\alpha$. Recalling that $u \in C_{o}$ implies that $u \not\in \mathrm{supp}(C_{o})$, as a consequence of~\eqref{pf:interpolator_bound_max_entry}, 
	\begin{equation*}
	\wt{p}_{\ell}(u) \leq \frac{\|\ell\|_1}{d_{\min}(\wt{G})}~~\textrm{for all $u \in C_{o}$,}
	\end{equation*}
	establishing~\eqref{eqn:ppr_uniform_bound_C}. The proof of~\eqref{eqn:ppr_uniform_bound_Cprime} follows similarly:
	\begin{equation*}
	p_v(u) = p_s(u) \overset{(i)}{\leq} \frac{\|s\|_1}{d_{\min}(C';G)} = \frac{\|\ell\|_1}{d_{\min}(C';G)},~~\textrm{for all $u \in C_{o}'$},
	\end{equation*}
	where the presence of $d_{\min}(C';G)$ on the right hand side of $(i)$ can be verified by inspecting~\eqref{pf:interpolator_bound_max_entry_inductive_step}. 	
\end{proof}

\subsection{Mixedness of lazy random walk and PPR vectors}
\label{subsec:lovasz_simonovits_bounds}
In this subsection, we give upper bounds on $h^{(t)} := h_{q_v^{(t)}}$ and $h^{(\alpha)} := h_{p_v}$. Although similar bounds exist in the literature (see in particular Theorem~1.1 of \citep{lovasz1990} and Theorem~3 of~\citep{andersen2006}), we could not find precisely the results we needed, and so for completeness we state and prove these results ourselves.

\begin{theorem}
	\label{thm:mixing_time_rw}
	For any $k \in [0, 2m]$, $t_0 \in \mathbb{N}$ and $t \geq t_0$,
	\begin{equation}
	\label{eqn:mixing_time_rw_1}
	h^{(t)}(k) \leq \frac{1}{2^{t_0}} + \frac{d_{\max}(G)}{d_{\min}(G)^2} + \frac{m}{d_{\min}(G)^2} \biggl(1 - \frac{\Psi(G)^2}{8}\biggr)^{t - t_0}.
	\end{equation}
\end{theorem}

\begin{theorem}
	\label{thm:mixing_time_PPR}
	Let $\phi$ be any constant in $[0,1]$. Either the following bound holds for any $t \in \mathbb{N}$ and any $k \in [d_{\max}(G),2m - d_{\max}(G)]$:
	\begin{equation*}
	h^{(\alpha)}(k) \leq \alpha t + \frac{2\alpha}{1 + \alpha} + \frac{d_{\max}(G)}{d_{\min}(G)^2} + \frac{m}{d_{\min}(G)^2} \left(1 - \frac{\phi^2}{8}\right)^{t},
	\end{equation*}
	or there exists some sweep cut $S_j$ of $p_v$ such that $\Phi(S_j;G) < \phi$.
\end{theorem}

The proofs of these upper bounds will be similar to each other (in places word-for-word alike), and will follow a similar approach and use similar notation to that of \citep{lovasz1990,andersen2006}. For $h: [0,2m] \to [0,1]$, $0 \leq K_0 \leq m$ and $k \in [K_0,2m - K_0]$, define
\begin{equation*}
L_{K_0}(k;h) = \frac{2m - K_0 - k}{2m - 2K_0}h(K_0) + \frac{k - K_0}{2m - 2K_0}h(2m - K_0)
\end{equation*}
to be the linear interpolant of $h(K_0)$ and $h(2m - K_0)$, and additionally let
\begin{equation*}
C(K_0;h) := \max\set{\frac{h(k) - L_{K_0}(k;h)}{\sqrt{\overline{k}}}: K_0 \leq k \leq  2m - K_0}.
\end{equation*}
where we use the notation $\overline{k} := \min\{k, 2m - k\}$, and treat $0/0$ as equal to $1$. Our first pair of Lemmas upper bound $h^{(t)}$ and $h^{(\alpha)}$ as a function of $L_{K_0}$ and $C(K_0)$. Lemma~\ref{lem:mixing_random_walk} implies that if $t$ is large relative to $\Psi(G)$, then $h^{(t)}(\cdot)$ must be small.

@@

\begin{lemma}[\textbf{c.f Theorem~3 of \citep{andersen2006}}]
	\label{lem:mixing_time_PPR}
	Let $\phi \in [0,1]$. Either the following bound holds for any $t \in \mathbb{N}$, any $K_0 \in [0,m]$, and any $k \in [K_0,2m - K_0]$:
	\begin{equation}
	\label{eqn:mixing_time_PPR}
	h^{(\alpha)}(k) \leq \alpha t + L_{K_0}(k; h^{(\alpha)}) + C(K_0;h^{(\alpha)})\sqrt{\overline{k}}\left(1 - \frac{\phi^2}{8}\right)^t
	\end{equation}
	or else there exists some sweep cut $S_{j}$ of $p_v$ such that $\Phi(S_j;G) < \phi$.
\end{lemma}
